---
title: "3.2.4 Solutions"
output: github_document
---

The following questions stem from 
[Section 3.2](http://r4ds.had.co.nz/data-visualisation.html#first-steps) of Chapter 3 (Data Visualization) in [R for Data Science](http://r4ds.had.co.nz/). The "solutions" offer some answers. Issues and pull requests are very welcome here.

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse) 
```

**1. Run `ggplot(data = mpg)`. What do you see?**

An empty graph, which appears as a grey box in the Plots tab of RStudio. You can think of this graph as a blank canvas that you can use as a foundation for plots.

**2. How many rows are in `mtcars`? How many columns?**

`str(mtcars)`, `nrow(mtcars)` in combination with `ncol(mtcars)`, and `dim(mtcars)` represent three ways you can arrive at the answer of 32 rows and 11 columns.

**3. What does the `drv` variable describe? Read the help `?mpg` to find out.**

`drv` describes whether a vehicle has front-, rear-, or four-wheel drive. These values are coded f, r, and 4, respectively.

**4. Make a scatterplot of `hwy` vs `cyl`.**

```{r fig.height = 4, fig.width = 4, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

**5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?**

```{r fig.height = 4, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

You get a scatterplot with categorical variables assigned to both the x and y axes. The scatterplot displays the combinations of class and drive train that exist in the data. This information isn't very helpful because you can't see how many vehicles fall within a given class-and-drive-train combination. 

To make the plot useful, you need to jitter the points with `geom_jitter`.

```{r  fig.height = 4, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, y = drv), width = 0.30, height = 0.30)
```