---
title: "3.3.1 Solutions"
output: github_document
---

The following questions stem from 
[Section 3.3](http://r4ds.had.co.nz/data-visualisation.html#aesthetic-mappings) of Chapter 3 (Data Visualization) in [R for Data Science](http://r4ds.had.co.nz/). The "solutions" offer some answers. Issues and pull requests are very welcome here.

### Load packages

```{r load-packages, message = FALSE}
library(tidyverse)
```

**1. What's gone wrong with this code? Why are the points not blue?**

```{r fig.height = 3.5, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The color aesthetic needs to be set outside of the aes function, as follows:

```{r echo = TRUE, fig.height = 3.5, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

**2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset.) How can you see this information when you run `mpg`?**

Categorical: `manufacturer`, `model`, `year`, `trans`, `drv`, `fl`, and `class`.

Continuous: `displ`, `cyl`, `cty`, and `hwy`.

```{r}
mpg
```

Running `mpg` displays the `mpg` data as a tibble. Under the variables' names, you can see their corresponding variable types, which are enclosed within arrows (e.g., below `manufacturer`, you see \<chr\>). Characters (chr), factors (fctr), and logicals (lgl) are all categorical variables, whereas integers (int) and floating point numbers (dbl) are both continuous variables.

**3. Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?**

For the `color` aesthetic, continuous variables get mapped along a color spectrum.
  
```{r fig.height = 4, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl))
```

Encoding `cyl`, the number of cylinders in a vehicle, as a categorical variable "fixes" this issue, such that vehicles with different numbers of cylinders get assigned to different colors.

```{r fig.height = 3.25, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = as.factor(cyl)))
```

For continuous variables that are discrete and have a small number values, the next plot shows that the `size` aesthetic behaves the same with continuous variables as it does with categorical variables.

```{r fig.height = 3.25, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cyl))
```

For the `shape` aesthetic, R returns an error message (i.e., "Error: A continuous variable cannot be mapped to shape").

```{r echo = TRUE, eval = F}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = cyl))
```

**4. What happens if you map the same variable to multiple aesthetics?**

Mapping a variable to the x or y axis as well as to another aesthetic creates redundancy. For example, in the next plot, the y axis and the legend provide redundant information.

```{r fig.height = 3.5, fig.width = 5, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```

**5. What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`.)**
  
The `stroke` aesthetic modifies the width of a data point's borders. It only works for shapes 21-24.

```{r fig.height = 3, fig.width = 4, fig.align = 'center'}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 22, stroke = .8)
```

**6. What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?**

`ggplot` uses logical arguments (like `aes(colour = displ < 5)`) to map values from data (like `displ == 6`) to levels of aesthetic properties (like `"red"`). In this case, the decision rule determines what color a data point gets assigned.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```